
In this blog, I want to perform Exploratory Data Analysis with Facebook dataset. This dataset contains almost 100,000 users and it varies from age, birthday,gender, to likes, mobile likes, etc..

<!-- TEASER_END -->

Well this isn't actually a real Facebook dataset. But this pseudo data is provided by Data  Analysts at Facebook. So  we can be assured it's as good as the real one. This Exploratory Data Analysis ranging from my experience from Udacity Course, Exploratory Data Analysis with R, in which I acquired the dataset. You should check it, it's really recommended course.


```{r}
pf <- read.csv('pseudo_facebook.tsv',sep='\t')
```

Here I generate the html using Knit HTML with Rstudio. the code is as given.

#Overview

Now, to get better at analyzing at the dataset, it's good to have all summary that we need to do this analysiz. First, I will do some basic summary to get better understand at the dataset. Here the dataset contain the words 'dob', which means data of birth.

```{r}
summary(pf)

head(pf)
```

#Female vs Male

As this dataset also contain the gender, I want to know every analysis that differentiate the male from the female. First let's take a look at each of the gender. Let's see the friend count for both female and male.

```{r}
by(pf$friend_count,pf$gender,summary)
```

The median of female is better than male, because the mean in female will drag the median lefwards of the graph. The median will resistance about the outliers(friend_count hight), because the average we can say that we at least try half of our data.

#Number of years using Facebook vs number of users in examples.

```{r}
library(ggplot2)
ggplot(aes(x = tenure/365), data =pf) + geom_histogram(binwidth = 0.25,color = 'black', fill = '#099DD9') + scale_x_continuous(breaks = seq(1,7,1) , limits = c(0,7)) + xlab('Number of years using Facebook') + ylab('Number of users in sample')

```

We see exponential growth of users in Facebook for the last 3 years. Adding the color, this graph will proceed as follows;

Tenure range from 1-365 days. in this case, I want to take range from 0 to 1 This will give me range histogram, to make it more convenience to see. Then I bulk(binwidth) it in 0.25.

Let's see users' age in this dataset.

```{r}
ggplot(aes(x = age), data =pf) + geom_histogram(binwidth=1) + scale_x_discrete(breaks=c(0,113,5))
```

For some people, they don’t have friend, so the baselog10 of zero(friend_count) = 0, this will get me negative towards infinity which will disrupt this data. So to avoid this I increment all friend count by 1. Natural Log will set better normal distribution over outliers in this data. I also using sqrt to compare log based 10 to convert it to normal distribution.

```{r}
library(gridExtra)
pl = ggplot(aes(x = friend_count), data = pf) + geom_histogram()
pl2 = pl + scale_x_log10()
pl3 = pl + scale_x_sqrt()
grid.arrange(pl,pl2,pl3,ncol=1)
```

#Female vs Male Continue

Now let's see which gender has likes by using web platform the most.

```{r}
ggplot(aes(x =www_likes, y = ..count../sum(..count..)),
       data = subset(pf, !is.na(gender))) + geom_freqpoly(aes(color = gender)) +
       scale_x_continuous(limits = c(0,500), breaks = seq(0,500,50))+
       xlab('count for www_likes') + ylab('ratio on which gender')
```

The data seems to tell me that female likes more often than male. Let's test this by using different approach.

```{r}
by(pf$www_likes,pf$gender,sum)
```

And by using boxplot,

```{r}
ggplot(aes(y = friend_count, x = gender), data = subset(pf, !is.na(gender))) + 
  geom_boxplot()  + scale_y_continuous(breaks = c(0,1000))
```

the boxplot specify that most of women has higher max friend_count than men


#On average, who initiated more friendships in our sample: men or women?

```{r}
ggplot(aes(y = friendships_initiated, x = gender), data = subset(pf, !is.na(gender))) + 
  geom_boxplot()  + coord_cartesian(ylim = c(0,150))

by(pf$friendships_initiated, pf$gender, summary)
```

Here I see that women initiated more friendship than men. Boxplot should have more advantage than mere by-function, and summary split each of gender, because it gives me better sense how determining the outlier of data, and also the normal distribution.

#Checking if the facebook mobile app is really benefit for the company

```{r}
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
sum(pf$mobile_check_in == 1)/length(pf$mobile_check_in)
```

With almost 65% user on mobile device, then I know why Facebook should make mobile app. It's very benefit for the company.

Don’t try to understand kinds of data that you have, but also what transformations can you made using your data. In this example I synthesize the data, rather than just mere naturallog/sqrt function



